$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-04-28 paper/paper.tex
Change title
Another title!
Itemize subject congruence proof

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Mechanising a linear \picalc}
\author{Uma Zalakain}{University of Glasgow, Scotland}
       {u.zalakain.1@research.gla.ac.uk}{https://orcid.org/0000-0002-3268-9338}{}
\author{Ornela Dardha}{University of Glasgow, Scotland}
       {ornela.dardha@glasgow.ac.uk}{https://orcid.org/0000-0001-9927-7875}{}
\authorrunning{U. Zalakain and O. Dardha}
\Copyright{Uma Zalakain and Ornela Dardha}
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003752.10003753.10003761.10003764</concept_id>
<concept_desc>Theory of computation~Process calculi</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}
\ccsdesc[300]{Theory of computation~Process calculi}
\keywords{pi-calculus, linear types, leftover typing, concurrency, mechanization, Agda}
\supplement{\url{https://github/umazalakain/typing-linear-pi}}
\acknowledgements{We want to thank Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

    \inferrule
    {l : \types{\gamma}{\Gamma}{P}{\Delta} \\\\
     r : \types{\gamma}{\Delta}{Q}{\Xi}}
    {\constr{comp} \; l \; r : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
  \end{mathpar}
  \caption{Left-over typing for a resource-aware typing system.}
  \label{fig:types}
  \end{figure}
\end{nidefinition}

@@

\begin{proof}
  By induction on \textsc{Equals}.
  The case $\constr{struct}$ and its symmetric one under $\constr{sym}$ are the nontrivial ones.
  For those, we proceed by induction on \textsc{StructCong}.
  The case $\constr{comp-assoc}$ is trivial because the encoding with leftover typing of the typing rules naturally preserves associativity.
  The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (Theorem \ref{thm:framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
  The middle-point of the resulting derivation on $\comp{Q}{P}$ will no longer be the same.
  The case $\constr{comp-end}$ is trivial because in the typing rule for process $\PO$ the input and output contexts are one and the same.
  In the $\constr{scope-end}$ case, the typing rules $\constr{chan}$ and $\constr{end}$ work together to constraint the multiplicity of the newly create channel to $\ell_{\o}$, making the proof trivial.
  In the opposite direction, we will need to instantiate the newly created channel to a concrete type and multiplicities.
  For this, we will use type $\unit$ and multiplicities $\ell_{\o}$.
  In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of Lemma $\ref{lm:types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
  We then use strengthening (Theorem \ref{thm:strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
  In the reverse direction, we will weaken $P$ so that it can accept an additional variable.
  We will then use Lemma \ref{lm:lower-lift} to show that lifting a lowered $P$ leaves $P$ unchanged.
  Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in Theorem \ref{thm:swapping}.
  In the reverse direction, we will use swapping and then use Lemma \ref{lm:swap-swap} to show that swapping indices twice leaves $P$ unchanged.
\end{proof}
\todo{make sure constructor names in code coincide}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{$\pi$ with leftovers: \\ a mechanisation in Agda}
\author{Uma Zalakain}{University of Glasgow, Scotland}
       {u.zalakain.1@research.gla.ac.uk}{https://orcid.org/0000-0002-3268-9338}{}
\author{Ornela Dardha}{University of Glasgow, Scotland}
       {ornela.dardha@glasgow.ac.uk}{https://orcid.org/0000-0001-9927-7875}{}
\authorrunning{U. Zalakain and O. Dardha}
\Copyright{Uma Zalakain and Ornela Dardha}
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003752.10003753.10003761.10003764</concept_id>
<concept_desc>Theory of computation~Process calculi</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}
\ccsdesc[300]{Theory of computation~Process calculi}
\keywords{pi-calculus, linear types, leftover typing, concurrency, mechanization, Agda}
\supplement{\url{https://github/umazalakain/typing-linear-pi}}
\acknowledgements{We want to thank Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

    \inferrule
    {l : \types{\gamma}{\Gamma}{P}{\Delta} \\\\
     r : \types{\gamma}{\Delta}{Q}{\Xi}}
    {\constr{comp} \; l \; r : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
  \end{mathpar}
  \caption{Leftover typing for a resource-aware typing system.}
  \label{fig:types}
  \end{figure}
  \todo{align typing rules to make them more readable}
\end{nidefinition}

@@

\begin{proof}
  By induction on \textsc{Equals}.
  The case $\constr{struct}$ and its symmetric one under $\constr{sym}$ are the nontrivial ones.
  For those, we proceed by induction on \textsc{StructCong}:
  \begin{itemize}
    \item
      The case $\constr{comp-assoc}$ is trivial because the encoding with leftover typing of the typing rules naturally preserves associativity.
    \item
      The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (Theorem \ref{thm:framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
      The usage context on which $Q$ and $P$ coincide in $\comp{Q}{P}$ will no longer be the same.
    \item
      The case $\constr{comp-end}$ is trivial because in the typing rule for process $\PO$ the input and output contexts are one and the same.
    \item
      In the $\constr{scope-end}$ case, the typing rules $\constr{chan}$ and $\constr{end}$ work together to constraint the multiplicity of the newly create channel to $\ell_{\o}$, making the proof trivial.
      In the opposite direction, we will need to instantiate the newly created channel to a concrete type and multiplicities.
      For this, we will use type $\unit$ and multiplicities $\ell_{\o}$.
    \item
      In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of Lemma $\ref{lm:types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
      We then use strengthening (Theorem \ref{thm:strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
      In the reverse direction, we will weaken $P$ so that it can accept an additional variable.
      We will then use Lemma \ref{lm:lower-lift} to show that lifting a lowered $P$ leaves $P$ unchanged.
    \item
      Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in Theorem \ref{thm:swapping}.
      In the reverse direction, we will use swapping and then use Lemma \ref{lm:swap-swap} to show that swapping indices twice leaves $P$ unchanged.
  \end{itemize}
\end{proof}
\todo{make sure constructor names in code coincide}
